humanCdGF_predict_cfu

### 2016-10-25 code/HumanCdGF_predict_cfu.Rmd
add cumulative importance for one mouse model

#One per cage
one_mouse_df <- data.frame(colonized_cfu %>% 
    right_join(community_d0.df) %>% 
    filter(!is.na(log_cfu_med))  %>% 
    select(log_cfu_med, mouse_id, cage_id, one_of(top_important_OTU)))

om_rsq_rf <- c() # rsq of each rf model
om_rsq <- c() # cumulative rsq of all models
om_import_otus <- data.frame(otu = colnames(select(one_mouse_df, contains('Otu'))))
one_df <- c()
om_pred <- c()
for(i in 1:20){
    set.seed(i)
    train <- one_mouse_df %>% 
      group_by(cage_id) %>% 
      sample_n(1) %>% ungroup() 
    test <- one_mouse_df %>% 
      filter(!mouse_id %in% train$mouse_id)
    set.seed(seed)
    rf_om <- randomForest(log_cfu_med ~ ., data = select(train, -mouse_id, -cage_id),
                             importance = TRUE, keep.forest = T, ntree = n_trees)
    om_rsq_rf[i] <- tail(rf_om$rsq, 1)
    om_pred <- rbind(om_pred,
                     data.frame(select(test, mouse_id, cage_id, log_cfu_med),
                          predicted=predict(rf_om, test)))
    om_rsq[i] <- 1 - (sum((om_pred$log_cfu_med - om_pred$predicted)^2)/
               sum((om_pred$log_cfu_med - mean(om_pred$log_cfu_med))^2))
    import_df <- importance(rf_om)
    colnames(import_df)[1] <- paste(train$mouse_id, collapse = ' ')
    om_import_otus <- merge(om_import_otus, import_df, by.x = 'otu', by.y = 'row.names')
    om_import_otus <- select(om_import_otus, -IncNodePurity)
}
plot(om_rsq_rf)
#####
test <- c() # test if any model has same train set
  for(i in 2:21){
  test[i-1] <- colnames(om_import_otus)[i] %in% colnames(om_import_otus)[-c(1,i)]
}
any(test)
#####

>>>>

#One per cage
one_mouse_df <- data.frame(colonized_cfu %>% 
    right_join(community_d0.df) %>% 
    filter(!is.na(log_cfu_med))  %>% 
    select(log_cfu_med, mouse_id, cage_id, contains('Otu')))

om_rsq_rf <- c() # rsq of each rf model
om_rsq <- c() # cumulative rsq of all models
om_import_otus <- data.frame(otu = colnames(select(one_mouse_df, contains('Otu'))))
one_df <- c()
om_pred <- c()
for(i in 1:20){
    set.seed(i)
    train <- one_mouse_df %>% 
      group_by(cage_id) %>% 
      sample_n(1) %>% ungroup() 
    test <- one_mouse_df %>% 
      filter(!mouse_id %in% train$mouse_id)
    set.seed(seed)
    rf_om <- randomForest(log_cfu_med ~ ., data = select(train, -mouse_id, -cage_id),
                             importance = TRUE, keep.forest = T, ntree = n_trees)
    om_rsq_rf[i] <- tail(rf_om$rsq, 1)
    om_pred <- rbind(om_pred,
                     data.frame(select(test, mouse_id, cage_id, log_cfu_med),
                          predicted=predict(rf_om, test)))
    om_rsq[i] <- 1 - (sum((om_pred$log_cfu_med - om_pred$predicted)^2)/
               sum((om_pred$log_cfu_med - mean(om_pred$log_cfu_med))^2))
    import_df <- importance(rf_om)
    colnames(import_df)[1] <- paste(train$mouse_id, collapse = ' ')
    om_import_otus <- merge(om_import_otus, import_df, by.x = 'otu', by.y = 'row.names')
    om_import_otus <- select(om_import_otus, -IncNodePurity)
}
test <- data.frame(om_import_otus[ , -1], row.names = om_import_otus$otu)
colnames(test) <- seq(ncol(test))

test_mean <- apply(test, 2, function(x) mean(x, na.rm=T)) # mean importance for each rf model
test_select <- apply(test, 1, function(x) x >= test_mean) # T for each otu above mean by sample
test_raw_mean <- apply(test_select, 2, function(x) sum(x[!is.na(x)])/length(x[!is.na(x)])) # portion of rf model otu is above mean imp
test_import_mean <- sort(test_raw_mean[test_raw_mean > 0.5], decreasing = T)

#### median with 
test_med <- apply(test, 2, function(x) median(x, na.rm=T)) # median importance for each rf model
test_select <- apply(test, 1, function(x) x >= test_med) # T for each otu above median by sample
test_raw_med <- apply(test_select, 2, function(x) sum(x[!is.na(x)])/length(x[!is.na(x)])) # portion of rf model otu is above med imp
test_import_med <- sort(test_raw_med[test_raw_med > 0.5], decreasing = T)

test_rank <- apply(test, 2, function(x) rank(x))
test_raw_rank <- apply(test_rank, 1, function(x) sum(x, na.rm = T))
sort(test_import_rank[test_raw_rank >=  ncol(test_rank)*sum(test_rank[ , 1])/nrow(test_rank)], decreasing = T)

#####
test <- c() # test if any model has same train set
  for(i in 2:21){
  test[i-1] <- colnames(om_import_otus)[i] %in% colnames(om_import_otus)[-c(1,i)]
}
any(test) # if true, then trained on the same set of mice multiple times
#####
